The \project{unRAVE} catalog

### 2016-08-18 article/unrave.tex
Data driven goodness

Given the high-quality of the normalization performed by the \rave\ team, we chose
not to re-normalize the spectra.  Our tests demonstrated that the procedure 
outlined in \citet{Kunder_2016} is sufficient for our analysis procedure. Therefore,
there is a limited number of pre-processing steps that we conducted before our
analysis could begin.  Firstly, we calculated inverse variance arrays from the
fractional uncertainties provided, then re-sampled the flux and inverse variance
arrays onto a common rest-wavelength map for all stars.  Depending on the fibre 
used and the stellar radial velocity, the range of rest-frame wavelength values
varied for each star.  Given that fluxes were unavailable in the edge pixels for 
most stars, we excluded pixels outside of the rest wavelength range 
$8423.2\,{\rm \AA} \le \lambda \le 8777.6\,{\rm \AA}$.  This corresponds to about
30~pixels excluded on either side of the common wavelength array.

@@

Given the relatively low S/N ratios in the \rave\ spectra, we chose to adopt a 
data-driven model (rather than a physics-driven model) for this analysis.

\stub{data-driven goodness}
% Describe The Cannon briefly, refer to Ness.

@@

We sought to construct a training set consisting of stars across the main-sequence,
sub-giant, and the red giant branch.  We require stars with precisely measured
effective temperature $\teff$, surface gravity $\logg$, and elemental abundances
of Fe, Mg, Si, Ti, Ca, Al, and Ni.  This proved to be difficult because the magnitude
range of \rave\ does not overlap substantially with high-resolution spectroscopic
surveys.  The fourth data release of the \project{Gaia-ESO} survey includes giant
and main-sequence stars, but only 142 overlap with \rave.  The most recent 
\project{APOGEE} data release \citep{sloan_dr13} reports labels for stars on the
giant branch and (uncalibrated values for) the main-sequence, but our tests indicated
that the main-sequence labels suffered from unexpected systematic effects.  A flat, 
then `up-turning' main-sequence is present, and the metallicity gradient trends in 
the opposite direction with respect to $\logg$ on the main-sequence (i.e., metal-poor
stars sit above the isochrones, not below).  If we move to lower-resolution studies,
there are 2,369 stars that overlap with \project{LAMOST} --- of which 2,213 have positive
S/N ratios in the $g$-band (\texttt{snrg}).  However, similar systematic effects are 
seen in the main-sequence, the labels are expectedly less precise given the lower
resolution, no detailed abundance information is available\footnote{Abundance 
information is available for \project{LAMOST} stars from \citet{Ho_2016}, but that 
sample contains only giant stars}, and there is only sparse sampling across the 
Hertsprung-Russell diagram.  The \project{GALAH} survey has a good overlap with \rave,
but results from that sample are not yet precise enough for our purposes.

@@

% Include velocities at test time?

>>>>

Given the high-quality of the normalization performed by the \rave\ team, we chose
not to re-normalize the spectra.  Our tests demonstrated that the procedure 
outlined in \citet{Kunder_2016} is sufficient for our analysis procedure. Therefore,
there is a limited number of pre-processing steps that we conducted before our
analysis could begin.  Firstly, we calculated inverse variance arrays from the
fractional uncertainties provided, then re-sampled the flux and inverse variance
arrays onto a common rest-wavelength map for all stars.  Depending on the fibre 
used and the stellar radial velocity, the range of rest-frame wavelength values
varied for each star.  Given that fluxes were unavailable in the edge pixels for 
most stars, we excluded pixels outside of the rest wavelength range 
$8423.2\,{\rm \AA} \le \lambda \le 8777.6\,{\rm \AA}$.  This corresponds to about
30~pixels excluded on either side of the common wavelength array, leaving us with
945~pixels for science.

@@

We chose to adopt a data-driven model --- rather than a physics-driven model ---
for this analysis.  Specifically we will use an implementation of \thecannon\
\citep{Ness_2015}.  Although this choice complicated the construction of our model
(e.g., see Section \ref{sec:the-training-set}), a data-driven model makes
use of all available information in the spectrum, and lowers the S/N ratio at
which systematic effects begin to dominate.  In other words, a well-constructed
data-driven model will yield more precise labels (e.g., stellar parameters and
chemical abundances) at low S/N ratios than most physics-driven models\footnote{
However, see \citet{Casey_2016}}.  This is particularly relevant for the 
low-resolution \rave\ data analysed here, because about half of the spectra have
S/N $\lesssim 50$.

Throughout this work we adopt the assumptions and nomenclature as set out in
\citet{Casey_2016}.  One distinction here is that the \emph{labelled} set and
the \emph{training} set in this work will be the same set.  In \citet{Casey_2016}
the labelled set contained two sub-sets: the training set and a distinct
validation set, allowing us to train on the \emph{training set} and then test
on the \emph{validation set}, a subset of the labelled set.  However in this work
the labelled set is relatively small, about an order of magnitude smaller than 
what was available in \citet{Casey_2016}, and therefore we will use all labelled 
set stars in the training step and supplement our validation experiments with 
literature comparisons.  Thus, throughout this work we will use the terms 
\emph{labelled set} and \emph{training set} interchangeably.  Below we describe 
how the training set was constructed before outlining the model we adopt.

@@

We sought to construct a training set consisting of stars across the main-sequence,
sub-giant, and the red giant branch.  We require stars with precisely measured
effective temperature $\teff$, surface gravity $\logg$, and elemental abundances
of Fe, Mg, Si, Ti, Ca, Al, and Ni.  This proved to be difficult because the magnitude
range of \rave\ does not overlap substantially with high-resolution spectroscopic
surveys.  The fourth internal data release of the \project{Gaia-ESO} survey includes 
giant and main-sequence stars, but only 142 overlap with \rave.  The most recent 
\project{APOGEE} data release \citep{sloan_dr13} reports labels for stars on the
giant branch and (uncalibrated values for) the main-sequence, but our tests indicated
that the main-sequence labels suffered from unexpected systematic effects.  A flat, 
then `up-turning' main-sequence is present, and the metallicity gradient trends in 
the opposite direction with respect to $\logg$ on the main-sequence (i.e., metal-poor
stars sit above the isochrones, not below).  If we move to lower-resolution studies,
there are 2,369 stars that overlap with \project{LAMOST} --- of which 2,213 have positive
S/N ratios in the $g$-band (\texttt{snrg}).  However, similar systematic effects are 
seen on the main-sequence: the labels are expectedly less precise given the lower
resolution, no detailed abundance information is available\footnote{Abundance 
information is available for \project{LAMOST} stars from \citet{Ho_2016}, but that 
sample contains only giant stars}, and there is only sparse sampling across the 
Hertsprung-Russell diagram.  The \project{GALAH} survey has a good overlap with \rave,
but results from that sample are not yet precise enough for our purposes.

% These restrictions forced us to use a heterogeneous training set.
%

\stub{We used the APOGEE DR13 cross-over for training set stars on the giant branch with cuts XYZ}

@@

% Include velocities and rotational velocities at test time?
